Understanding the Lomb-Scargle Periodogram

### 2017-02-02 PracticalLombScargle.tex
some rewordings

\subsection{Outline}
The paper is organized as follows:
In \sect{continuous-fourier-transform}, we review the continous Fourier
transform and some of its useful properties, including defining the notion of
a power spectrum for measuring periodic content in a signal.
In \sect{window-functions}, we use these properties to understand how different
observation patterns (i.e. window functions) affect the power spectrum, and
discuss a conceptual understanding of the well-known Nyquist limiting frequency.
In \sect{non-uniform-sampling}, we consider non-uniformly sampled signals as
a special case of such windowing, and show that the Nyquist-like limit for
this case is quite different than what many practitioners in the field
assume.
In \sect{schuster-to-lomb-scargle}, we discuss the properties of the
Lomb-Scargle periodogram, a modified version of the classical periodogram
for unevenly-sampled data, as well as the motivation behind these modifications.
In \sect{lomb-scargle-extensions}, we discuss a complementary view of the
periodogram as the result of least-squares model fitting, as well as some
extensions that allows (including the ``Bayesian'' periodogram).
In \sect{practical-considerations}, we discuss important practical
considerations for the use of the Lomb-Scargle periodogram, including
the choice of frequency grid, false alarm probabilities, and various modes
of failure that should be expected when working with real-world data.
Finally in \sect{conclusion}, we conclude and summarize the key recommendations
for practical use of the Lomb-Scargle periodogram.

\subsection{Code and Figures}
All figures in this paper were produced using Python, and in particular the
Numpy \citep{numpy, numpybook},
Pandas \citep{pandas},
Astropy \citep{Astropy2013},
and Matplotlib \citep{matplotlib} packages.
Code to reproduce these figures is available in Jupyter notebooks at
\url{http://github.com/jakevdp/PracticalLombScargle/}.

@@

\section{Practical Considerations when using Lomb-Scargle Periodograms}
\sectlabel{practical-considerations}

@@

All periodograms in this paper were computed using the Astropy
implementations of the \citet{Press89}, \citet{Zechmeister09}, and
\citet{Palmer09} algorithms, which were adapted from Python code originally
published by \citet{ICVG2014} and \citet{VanderPlas2015}.
All code behind this paper is available in the form of Jupyter notebooks in the
paper's GitHub repository\footnote{\url{http://github.com/jakevdp/PracticalLombScargle}}.

@@

One point of confusion in periodogram implementations is the interpretation
of the periodogram values and their meaning.
Recall that there are two equally valid perspectives of what the periodogram
measures: the Fourier view and the least-squares view.
Each of these suggests a different natural approach to normalizing and interpreting the periodogram.

@@

While the normalization does not affect the shape of the periodogram,
its main practical consequence is that the statistics of the resulting
periodogram differ slightly, and this needs to be taken into account when
computing analytic estimates of uncertainties and false alarm probabilities.
Other normalizations exist as well, but seem to be rarely used in practice.
For a concise summary of periodogram normalizations and their statistical
properties, refer to the introduction of \citet{Baluev2008}.

@@

This can be proven mathematically \citep[see, e.g.][]{Bretthorst88, ICVG2014},
but it is perhaps more intuitive to see this visually.
\fig{peak-width-height} shows periodograms for simulated data with a fixed
baseline $T=4$ units covering four cycles, with varying sample sizes and
signal-to-noise ratios.
In all cases, the width of the peaks (most clearly visible in one or another
normalization) are essentially {\it identical}, regardless of the data
quality or quantity!
Evidently, peak width ({i.e.} frequency precision) is not a particularly
useful measure of uncertainty in the detection of periods.

@@

While the False Alarm Probability can be a useful concept, you should be
careful to keep in mind that it answers a very specific question:
{\it what is the probability that a signal with no periodic component would
lead to a peak of this magnitude?}
In other words, it is a measure of the false positive rate: the rate at which
we would expect to incorrectly identify a periodic signal where none
actually exists.
It tells us nothing about the false negative rate; {i.e.} the rate at which
we would expect to miss a periodic signal that is, in fact, present.
It also tells us nothing about the error rate; {i.e.} the rate at which we
would expect to identify an incorrect alias of a signal present in our data
(as we saw in \sect{failure-modes}).
Too often users are tempted to interpret the FAP too broadly, with the hope that
it would speak to these additional questions as well.

To the best of my knowledge, there is currently no silver bullet for answering
these broader questions within the Lomb-Scargle approach.
Perhaps the most fruitful path toward understanding of such effects for a
particular survey---with particular noise characteristics and a particular
observing window---is via simulated data injected into the detection
pipeline.
In fact, this type of simulation is a vital component of most (if not all)
current and future time-domain surveys that seek to catalog periodic objects,
regardless of the particular approach used to detect periodicity
\citep[see, e.g.][]{opsim1, opsim2, Ivezic08LSST, Sesar2010, Oluseyi12, McQuillan2012, Drake2014}

>>>>

\subsection{Outline}
The paper is organized as follows:
\begin{description}
  \item[\sect{introduction}] is this introduction.
  \item[\sect{continuous-fourier-transform}] presents a review of the 
    continous Fourier transform and some of its useful properties,
    including defining the notion of a power spectrum for measuring
    periodic content in a signal.
  \item[\sect{window-functions}] builds on these properties to explore how
    different observation patterns (i.e. window functions) affect the observed
    power spectrum, and discusses a conceptual understanding of the
    well-known Nyquist limiting frequency.
  \item[\sect{non-uniform-sampling}] considers non-uniformly sampled signals as
    a special case of an observational window, and shows that the Nyquist-like
    limit for this case is quite different than what many practitioners in 
    the field often assume.
  \item[\sect{schuster-to-lomb-scargle}] details the Lomb-Scargle periodogram,
    a modified version of the classical periodogram for unevenly-sampled data,
    as well as the motivation behind these modifications.
  \item[\sect{lomb-scargle-extensions}] discusses a complementary view of the
    Lomb-Scargle periodogram as the result of least-squares model fitting,
    as well as some extensions enabled by that viewpoint
    (including various ``Bayesian'' periodograms).
  \item[\sect{practical-considerations}] utilizes the concepts introduced
    in earlier sections to discuss important practical considerations for
    the use of the Lomb-Scargle periodogram, including the choice of
    frequency grid, uncertainties and false alarm probabilities, and modes of
    failure that should be expected when working with real-world data.
  \item[\sect{conclusion}] concludes and summarizes the key recommendations
    for practical use of the Lomb-Scargle periodogram.
\end{description}
All figures in this paper were produced using Python, and in particular the
Numpy \citep{numpy, numpybook},
Pandas \citep{pandas},
Astropy \citep{Astropy2013},
and Matplotlib \citep{matplotlib} packages.
Code to reproduce these figures is available in Jupyter notebooks at
\url{http://github.com/jakevdp/PracticalLombScargle/}.

@@

\section{Practical Considerations when using Lomb-Scargle Periodograms}
\sectlabel{practical-considerations}
The previous sections have given a conceptual introduction to the Lomb-Scargle
periodogram, intertwining the Fourier and Least-squares views.
With these principles understood, this section gets to the meat of the subject
at hand: given this understanding of the Lomb-Scargle periodogram and related
approaches, what practical issues are involved in its effective use?

@@

All periodograms in this paper were computed using the Astropy
implementations of the \citet{Press89}, \citet{Zechmeister09}, and
\citet{Palmer09} algorithms, which were adapted from Python code originally
published by \citet{ICVG2014} and \citet{VanderPlas2015}.
All code behind this paper is available in the form of Jupyter notebooks in the
paper's GitHub repository at \url{http://github.com/jakevdp/PracticalLombScargle/}.

@@

Often it is useful to be able to interpret the periodogram values themselves.
Recall that there are two equally valid perspectives of what the periodogram
is measuring---the Fourier view and the least-squares view---and each of these
lends itself to a different approach to normalizion and interpretation of the
periodogram.

@@

While the normalization does not affect the shape of the periodogram,
its main practical consequence is that the statistics of the resulting
periodogram differ slightly, and this needs to be taken into account when
computing analytic estimates of uncertainties and false alarm probabilities
(see the next section).
Other normalizations exist as well, but seem to be rarely used in practice.
For a concise summary of periodogram normalizations and their statistical
properties, refer to the introduction of \citet{Baluev2008}.

@@

This can be proven mathematically \citep[see, e.g.][]{Bretthorst88, ICVG2014},
but it is perhaps more intuitive to see this visually.
\fig{peak-width-height} shows periodograms for simulated data with a fixed
baseline $T=4$ units covering four cycles, with varying sample sizes and
signal-to-noise ratios.
In all cases, the width of the peaks (most clearly visible in one or another
normalization) are essentially {\it identical}, regardless of the quality
or quantity of data!
Evidently, peak width ({i.e.} frequency precision) is not a particularly
useful measure of uncertainty in the detection of periods.

@@

While the False Alarm Probability can be a useful concept, you should be
careful to keep in mind that it answers a very specific question:
{\it ``what is the probability that a signal with no periodic component would
  lead to a peak of this magnitude?''}.
In particular, it emphatically does {\it not} answer the much more relevant
question {\it ``what is the probability that this data is periodic
  given these observations?''.}
This boils down to a case of $P(A~|~B) \ne P(B~|~A)$, but still it is common to
see the false alarm probability treated as if it speaks to the second rather
than the first question.
In other words, the false alarm probability tells you the false positive rate
in a frequentist, rather than Bayesian sense: that is, it is a false positive
rate conditioned on the entire space of possible observations rather than on
the particular data you observed.

Similarly, the false alarm probability tells us nothing about the false
negative rate; {i.e.} the rate at which
we would expect to miss a periodic signal that is, in fact, present.
It also tells us nothing about the error rate; {i.e.} the rate at which we
would expect to identify an incorrect alias of a signal present in our data
(as we saw in \sect{failure-modes}).
Too often users are tempted to interpret the FAP too broadly, with the hope that
it would speak to questions that are beyond its reach.

To the best of my knowledge, there is currently no silver bullet for answering
these broader, more relevant questions of uncertainty of Lomb-Scargle results.
Perhaps the most fruitful path toward understanding of such effects for a
particular set of observations---with particular noise characteristics and a
particular observing window---is via simulated data injected into the detection
pipeline.
In fact, this type of simulation is a vital component of most---I might even
venture to say all---current and future time-domain surveys that seek to
detect, characterize, and catalog periodic objects,
regardless of the particular approach used to identify periodicity
\citep[see, e.g.][]{opsim1, opsim2, Ivezic08LSST, Sesar2010,
  Oluseyi12, McQuillan2012, Drake2014}
